---
title: "Malaria"
author: "Zachary Hamilton"
date: "20/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

```{r import and clean}
library(tidyverse)
library(gapminder)
library(wbstats)
theme_set(theme_light())

malaria_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-11-13/malaria_deaths_age.csv")


malaria_processed <- malaria_raw %>%
  select(-X1)

pop_data <- wb(indicator = "SP.POP.TOTL", startdate = 2016, enddate = 2016) 

malaria_per_cap <- malaria_processed %>%
  left_join(select(pop_data, iso3c, value), by = c("code" = "iso3c")) %>%
  mutate(deaths_per_100k = deaths/value * 100000) %>%
  filter(year == 2016, !is.na(code), entity != "World")
```


```{r deaths over time by age }

malaria_processed %>%
  group_by(year, age_group) %>%
  summarise(mean(deaths)) %>%
  ggplot(aes(year, `mean(deaths)`, col = as.factor(age_group))) +
  geom_line(size = 1.1) +
  geom_point(size = 2) +
  labs(title = "Malaria Deaths by age group over time",
       x = 'Year', 
       y = 'Dpaths',
       col = 'Age Range')

```


The Under 5 Age group stands out as the demographic that is most often fatally affected by malaria.

Let's look on a map to see what regions of the world are most affected in 2016, our most current data, in this age group 

```{r top countries under 5}

malaria_processed %>%
  filter(year == 2016, age_group == 'Under 5') %>%
  mutate(entity = fct_reorder(entity, deaths)) %>%
  filter(deaths > 0) %>%
  arrange(desc(deaths)) %>%
  filter(!is.na(code), entity != 'World') %>%
  head(15) %>%
  ggplot(aes(entity, deaths, fill = entity)) +
  geom_col() + 
  coord_flip() +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  labs(title = "Countries affected by malaria in 2016",
       subtitle = "Only Under 5 Age Group Shown",
        x = "Country")
  

```

Nigeria had roughly triple the amount of deaths as the next highest country, the overwhelming majority are African countries, with the major exception of India, which makes me think I should probably make some sort of per capita adjustment instead of looking at overall counts. Let's find a world population table to join on and make some calculations - good old gapminder might do the trick. 

```{r gapminder test}

gapminder %>%
  summarize(max(year))

```

Most recent data available in gapminder dataset is 2007, this most certainly WILL NOT do! 

```{r per capita}
malaria_processed %>%
  glimpse()

pop_data <- wb(indicator = "SP.POP.TOTL", startdate = 2016, enddate = 2016) 

malaria_processed %>%
  left_join(select(pop_data, iso3c, value), by = c("code" = "iso3c")) %>%
  mutate(deaths_per_100k = deaths/value * 100000) %>%
  filter(year == 2016, age_group == 'Under 5') %>%
  mutate(entity = fct_reorder(entity, deaths_per_100k)) %>%
  filter(deaths > 0) %>%
  filter(!is.na(code), entity != 'World') %>%
  arrange(desc(deaths_per_100k)) %>%
  head(15) %>%
  ggplot(aes(entity, deaths_per_100k, fill = entity)) +
  geom_col() + 
  coord_flip() +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  labs(title = "Countries affected by malaria in 2016",
       subtitle = "Only Under 5 Age Group Shown",
        x = "Country")
  

```

poof, just like that India disappears from our chart because of its massive population rescaling factor. Burkina Faso, which has a population of `r as.double(filter(pop_data, country == "Burkina Faso")["value"])` Lets look at this world distribution on a map. 

```{r}
malaria_per_cap %>%
  glimpse()

world_map <- map_data("world") %>%
  mutate(code = countrycode::countrycode(region, "country.name", "iso3c"))
  
world_map %>%
  left_join(select(malaria_per_cap, code, entity, value, age_group, deaths, deaths_per_100k), by = "code") %>%
  filter(age_group != 'Under 5', !is.na(age_group)) %>%
  ggplot(aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = deaths_per_100k)) +
  lims(y = c(-55, 90)) +
  scale_fill_viridis_c(end = .75, option = "B") +
  theme(axis.line = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      panel.background = element_blank(),
      panel.border = element_blank(),
      panel.grid = element_blank(),
      panel.spacing = unit(0, "lines"),
      plot.background = element_blank(),
      legend.justification = c(0, 0),
      legend.position = c(0, 0)) +
  facet_wrap(~ age_group)
```

