---
title: "tidy_tuesday_2018-04-09"
author: "Zachary Hamilton"
date: "28/10/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prework, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(scales)

nfl_salary <- readxl::read_xlsx(here("data", "nfl_salary.xlsx"))

theme_set(new = theme_light())


```

Data set appears to be NFL salaries for various positions from 2011 to 2018. Would rather have year, position, salary columns. Good job for gather function. 

```{r cleaning, echo=FALSE}
nfl_salary_tidy <- nfl_salary %>%
  gather(key = "position", value = "salary", 2:11) %>%
  filter(!is.na(salary))
```


# First Question that jumps out in my mind is what position makes the most money?

```{r}
nfl_salary_tidy %>%
  mutate(position = fct_reorder(position, salary, median)) %>%
  ggplot(aes(x = position, y = log(salary))) +
  geom_boxplot() +
  coord_flip() 
  
```

Linemen seem to be consistently near the top in median salary each year. Most variation in the quarterback position. Consistently have highest paid + lowest paid players in the NFL. in 2018,9 of top 10 paid players were quarterbacks, as were 9 of 10 lowest paid players! This is likely because teams carry 3rd string quarterbacks who often won't take a snap for the entire season, whereas there are multiple starting offensive & defensive lineman who play an integral role to their teams success each game. 

```{r}
nfl_salary_tidy %>%
  filter(year == 2018) %>%
  group_by(year) %>%
  arrange(salary)
```

Has the leagues average salary gone up?

```{r}
nfl_salary_tidy %>%
  group_by(year, position) %>%
  summarize(median = median(salary)) %>%
  ggplot(aes(year, median, col = position)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = dollar_format())
```

